Draft Description of Work in Progress

### 2019-01-31 Naive_bayes_V2yn.Rmd
Working text analysis forms

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

@@

```{}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

#Rename the columns.
names(comments_rawVERB)[1] <- "subject"
names(comments_rawVERB)[2] <- "structure"
names(comments_rawVERB)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVERB$subject <- factor(comments_rawVERB$subject)
str(comments_rawVERB$subject)
table(comments_rawVERB$subject)
```

```{r}
#Change "structure" element from character to a factor for analysis.
comments_rawVERB$structure <- factor(comments_rawVERB$structure)
str(comments_rawVERB$structure)
table(comments_rawVERB$structure)
```

```{r}
#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)
```

```{r}
#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVERB$randomizer <- c(vector)
comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
print(comments_rawVERB_R)

```

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB_R[[5]])
```

@@

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())

#The option to use stemming has been removed from this version.

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

@@

This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

Make sure that the proportion of each sub-category is similar in the training and testing data sets.
```{r}
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))
```

```{r}
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
```

@@

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
```

###Train Model, Predict, Evaluate for structure
This is the same block as above but uses the "structure" data column.

```{r}
comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for structure.

```{r}
print("Overall structure accuracy")
(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{}
predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{}
predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

***
START COMMENTS, OLD work notes
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.

Permutations to test with random subsets version (2.0.r):

*  Compare 1-grams, 2-grams, 3-grams
*  Compare outcomes for 3 independent randomized sets of 1-grams

The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix. The second problem is how to pull the training and test data sets out and process them identically.

```{}
set.seed(1234) 		# ensures the same random set gets picked each time.

comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)

comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
```

The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
1. Create a vector of sequential numbers the same length as the number of data frame rows.
2. Randomize the sequential numbers using set.seed to make pattern reproducible.
3. Add the randomized numbers as a new column to the data frame.
4. Re-order the data table to the randomizer column.

This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.

```{}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

```{}
if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

#glimpse(tally)

#tally %>% mutate(match_subject=comments_test_labels_subject - comments_test_pred_subject)
#tally %>% mutate(match_structure=comments_test_labels_structure - comments_test_pred_structure)
```

>>>>

---
title: "Naive Bayes Workflow Ver 2.0.yn"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This is **Version 2.0.yn** of the basic workflow for classifying TA comments using a naive Bayes model. It is optimized for testing pair-wise assignment to categories (Yes/No decision) beginning with the most frequently encountered.

The main features and permutations to be evaluated are:

*  Stepwise paired classification versus comparing multiple groups simultaneously

Additional working notes are located at the end of this page.

@@

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0.yn of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

#Rename the columns.
names(comments_rawVERB)[1] <- "type"
names(comments_rawVERB)[2] <- "locus"
names(comments_rawVERB)[3] <- "text"

#Change "type" element from character to a factor for analysis.
comments_rawVERB$type <- factor(comments_rawVERB$type)
str(comments_rawVERB$type)
table(comments_rawVERB$type)
```

```{r}
#Change "locus" element from character to a factor for analysis.
comments_rawVERB$locus <- factor(comments_rawVERB$locus)
str(comments_rawVERB$locus)
table(comments_rawVERB$locus)
```

***
***
Here is where I need to add the column used to randomly assign the data to training or testing sets.

Use set.seed(1234) to create a stable randomized set.

***
***

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB <- VCorpus(VectorSource(comments_rawVERB$text))
print(comments_corpusVERB)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB[[5]])
```

@@

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB <- tm_map(comments_corpusVERB, content_transformer(tolower))
```

```{r}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removePunctuation)
```

```{r}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeNumbers)
```

```{r}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeWords, stopwords())
```

Option to stem terms has been removed from this version.

Final prep and data check.
```{r}
#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB[[5]]))
```

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

@@

This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
```{}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{}
comments_train_labels_type <- comments_rawVERB[1:7005, ]$type
comments_test_labels_type <- comments_rawVERB[7006:9340,]$type
```

```{}
comments_train_labels_locus <- comments_rawVERB[1:7005, ]$locus
comments_test_labels_locus <- comments_rawVERB[7006:9340,]$locus
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{}
prop.table(table(comments_train_labels_type))
prop.table(table(comments_test_labels_type))
```

```{}
prop.table(table(comments_train_labels_locus))
prop.table(table(comments_test_labels_locus))
```

@@

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no.

```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

@@

###Train Model, Predict, Evaluate for Type
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
comments_test_pred_type <- predict(comments_classifier_type, comments_test)
CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.

```{r}
print("Overall TYPE accuracy")
(14+45+1346+19)/23.35
```

###Train Model, Predict, Evaluate for LOCUS
This is the same block as above but uses the "Locus" data column.

```{r}
comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.

```{r}
print("Overall LOCUS accuracy")
(62+64+931+3)/23.35
```

@@

```{}
predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
head(predicted_category_type)
write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
```

```{}
predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
head(predicted_category_locus)
write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
```

***
START COMMENTS
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.

Permutations to Test:

*  Compare 1-grams, 2-grams, 3-grams
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Initial approach will be to split between largest group and balance in each category
    +  Ex.: type is 3-technical vs. not 3_technical

Known Issues:

1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.

### 2019-01-31 VERB.Rmd
Working text analysis forms

##Deeper Text Analysis
These linked pages summarize work to date on each method.

*  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
*  Random Forest (link)
*  Latent Dirichlet Allocation (link)
*  Etc. (link)
*  Etc. (link)

@@

#Initial Dataset
My working dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

#Data Pre-Checks
Prior to import, review data table in Excel. Check:

1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.rmd"

@@

```{r}
#This block reads in the full CSV into a starting dataframe named "base_data".
#Subsequent analyses call in subsets of "base_data".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

@@

@@

```{r}
# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)

@@

####Comparing Pearson Correlations of Word Frequencies
If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
```{r}
#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
#Stats need this format to be able to compare 2. WQ against other values.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)

@@

```{}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

>>>>

##Deeper Text Analysis
These linked pages summarize work to date on individual methods.

*  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
*  To Be Decided #1. (link)
*  To Be Decided #2. (link)
*  To Be Decided #3. (link)
*  Etc. (link)

@@

#Initial Dataset
The development dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

#Data Pre-Checks
Prior to import, review data table **in Excel**. Check:

1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"

@@

```{r}
#This block reads in the full CSV into a starting dataframe named "base_data".
#Subsequent analyses call in subsets of "base_data".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')

@@

For recoding comment labels, I am considering recoding **comment.subject** as:
```
0_null
1_basic
2_writing
3_technical
4_logic
5_exposition
6_misconduct
7_administrative
8_nobasis
9_split
10_context
11_flag
12_narrative
13_sciname
```
Recoding **comment.structure** as:
```
0_null
1_pointer
2_copyedit
3_general
4_specific
5_rationale
6_holistic
7_idiomatic
8_nobasis
```
Any obvious problems with these?

@@

```{r}
# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing <- filter(base_data, code.subject=="2_Writing"|code.subject=="3_Technical"|code.subject=="4_Logic")
frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)

@@

####Comparing Pearson Correlations of Word Frequencies
If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall.
```{r}
#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
#Stats need this format to be able to compare 2. WQ against other values.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)

@@

```{r}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
